Multi-Agent Pathfinding with Matching using Increasing Cost Tree Search

### 2021-06-26 docs/paper.tex
Finished rework of exhaustive ICTS

			\EndProcedure
		\end{algorithmic}
		\caption{Increasing Cost Tree Search}
		\label{icts-algo}
	\end{algorithm}
	\subsection{Exhaustive ICTS}
	\label{exhaustive}
	For $K$ teams $t_1,\ldots,t_K$ of size $k_i$, there are in total $M = \prod_{i} k_i!$ matchings from team agents to team goals. These matchings can be enumerated by indexing the $k_i!$ permutations of goals for each team $t_i$ as $1,\ldots,k_i!$. Using such indexing, each $e \in \{1,\ldots,k_1!\} \times \cdots \times \{1,\ldots,k_K!\}$ corresponds to a unique combination of permutations and can be mapped to a matching $m$ assuming some fixed ordering of agents within each team. Let $C_1$ through $C_M$ denote the optimal SIC costs of each MAPF instance corresponding to the MAPFM instance, with $C_m = \infty$ for infeasible matching $m$. In order to find the optimal solution, $m$ with lowest $C_m$ has to be found, which requires all $M$ matchings to be solved or found to be infeasible, meaning that $\mathcal{O}(M)$ MAPF instances have to be solved. 
	\subsubsection{Bounded search}
	Once a solution is found for a matching $m$ with cost $C_m$, this can be used in the next search as an (exclusive) upper-bound $B$ for the SIC of nodes that have to be searched. Formally, let $B_0 = \infty$ represent the initial SIC bound of the ICT search. After solving each $m$ with solution cost $C_m$, let $B_{m+1} = \min(B_m,C_m)$. By successively updating this $B$ in this manner, as more matchings are solved, the effort needed to solve the remaining matchings can be expected to be reduced significantly. With $C_m$ denoting the optimal SIC for matching $m$, this effort can be expressed as	
	\[\Delta_{m,\text{bounded}} = \min(B_m,C_m) - C^*_m = B_{m+1} - C^*_m\]
	This is proportional to the effort due to the complexity of the ICTS search for matching $m$, $\mathcal{O}(k^{\Delta_{m,\text{bounded}}})$.
	If $B_m < C^*_m$, only the root node will be generated to determine this so $\Delta_{m,\text{bounded}} = 0$.
	Using this, the number of ICT nodes searched in total becomes $\mathcal{O}(\sum_m k^{\Delta_{m,\text{bounded}}})$.
	\subsubsection{Ordered enumeration}
	\label{ordered-enum}
	To reduce the total effort, the matchings $m$ can instead be enumerated and solved in increasing order of $C^*_m$. In this case, the effort per matching is monotonically non-increasing by Property \ref{monotonicity}. This induces a case in which this enumeration minimizes the number of nodes searched in total (Property \ref{cond-monoton-optim}). However, cases in which $\Delta_m$ of the first $m$'s searched is much higher could also be constructed, in which case this approach might cause more nodes to be processed in total than alternative methods. Therefore, this ordered enumeration remains a heuristic improvement that works well in practice, as shown in Section \ref{experiments}.

	\begin{property}[Monotonicity]
		\label{monotonicity}
		In bounded exhaustive search using ICTS, solving the matchings in increasing order of root SIC makes the effort $\Delta_{m,\text{bounded}}$ monotonically non-increasing as a function of $m$.
	\end{property}
	\begin{proof}
		Due to the order of matchings for successive matchings $m,m+t$ for some $t\geq 1$, $C^*_{m} \leq C^*_{m+t}$. By definition, $B_{m+2} = \min(B_{m+1},C_{m+1})$ meaning that $B_{m+2} \leq B_{m + 1}$ so transitively $B_{m+t+1} \leq B_{m+1}$. Therefore
		\[B_{m+t+1} - C^*_{m+t}\leq B_{m+1} - C^*_m\implies \Delta_{m+t,\text{bounded}} \leq \Delta_{m,\text{bounded}}\]
	\end{proof}
	\begin{property}[Conditional optimality]
		\label{cond-monoton-optim}
		Assuming that $\Delta_m$, the unbounded effort, is monotonically non-decreasing as a function of $m$ (with $m$ in increasing order of $C^*_m$), ordered enumeration minimizes the total number of nodes searched across matchings.
	\end{property}
	\subsubsection{Approximate ordered enumeration}
	In implementation matchings were enumerated using a product of goal permutation indices as described above so they are not enumerated in increasing order of root SIC. As a result, to achieve this ordering, matchings have to be exhaustively enumerated and stored in memory in order to be sorted. This requires $\mathcal{O}(M)$ memory which was found to be more problematic in practice than the time complexity of the sorting, $\mathcal{O}(M\log(M))$.

	%Using the 
	%To generate matchings 
	%In the implementation of exhaustive matching, the algorithm used to generate matchings
	%The algorithm for generating 
	%In the implementation of ordered enumeration as described above, matchings are not generated in increasing order of root SIC. As a result, to achieve this ordering, matchings have to be exhaustively enumerated and stored in memory in order to be sorted. This requires $\mathcal{O}(M)$ memory which was found to be more problematic in practice than the time complexity of the sorting, $\mathcal{O}(M\log(M))$.

@@

	This is why in implementation, ordered enumeration is approximated using a priority queue that maintains a subset of matchings of a fixed size $M_{pq}$\footnote{$M_{pq} = 10000$ was used in implementation but this could be much increased at the cost of memory}. Algorithm \ref{pq-algo} outlines the procedure.

>>>>

			\EndProcedure
		\end{algorithmic}
		\caption{Increasing Cost Tree Search}
		\label{icts-algo}
	\end{algorithm}
	\subsection{Exhaustive ICTS}
	\label{exhaustive}
	For $K$ teams $t_1,\ldots,t_K$ of size $k_i$, there are in total $M = \prod_{i} k_i!$ matchings from team agents to team goals. These matchings can be enumerated by indexing the $k_i!$ permutations of goals for each team $t_i$ as $1,\ldots,k_i!$. Using such indexing, each $e \in \{1,\ldots,k_1!\} \times \cdots \times \{1,\ldots,k_K!\}$ corresponds to a unique combination of permutations and can be mapped to a matching $m$ assuming some fixed ordering of agents within each team. Let $C_1$ through $C_M$ denote the optimal SIC costs of each MAPF instance corresponding to the MAPFM instance, with $C_m = \infty$ for infeasible matching $m$. In order to find the optimal solution, $m$ with lowest $C_m$ has to be found, which requires all $M$ matchings to be solved or found to be infeasible, meaning that $\mathcal{O}(M)$ MAPF instances have to be solved. 
	\subsubsection{Bounded search}
	Once a solution is found for a matching $m$ with cost $C_m$, this can be used in the next search as an (exclusive) upper-bound $B$ for the SIC of nodes that have to be searched. Formally, let $B_0 = \infty$ represent the initial SIC bound of the ICT search. After solving each $m$ with solution cost $C_m$, let $B_{m+1} = \min(B_m,C_m)$. By successively updating this $B$ in this manner, as more matchings are solved, the effort needed to solve the remaining matchings can be expected to be reduced significantly. With $C_m$ denoting the optimal SIC for matching $m$, this effort can be expressed as	
	\[\Delta_{m,\text{bounded}} = \min(B_m,C_m) - C^*_m = B_{m+1} - C^*_m\]
	This is proportional to the effort due to the complexity of the ICTS search for matching $m$, $\mathcal{O}(k^{\Delta_{m,\text{bounded}}})$.
	If $B_m \leq C^*_m$, only the root node will be generated to determine this so $\Delta_{m,\text{bounded}} = 0$.
	Using this, the number of ICT nodes searched in total becomes $\mathcal{O}(\sum_m k^{\Delta_{m,\text{bounded}}})$.
	\subsubsection{Ordered enumeration}
	\label{ordered-enum}
	To reduce the total effort, the matchings $m$ can instead be enumerated and solved in increasing order of $C^*_m$. In this case, the effort per matching is monotonically non-increasing by Property \ref{monotonicity}. This induces a case in which this enumeration minimizes the number of nodes searched in total (Property \ref{cond-monoton-optim}). However, cases in which $\Delta_m$ of the first $m$'s searched is much higher could also be constructed, in which case this approach might cause more nodes to be processed in total than alternative methods. Therefore, this ordered enumeration remains a heuristic improvement that works well in practice, as shown in Section \ref{experiments}.

	\begin{property}[Monotonicity]
		\label{monotonicity}
		In bounded exhaustive search using ICTS, solving the matchings in increasing order of root SIC makes the effort $\Delta_{m,\text{bounded}}$ monotonically non-increasing as a function of $m$.
	\end{property}
	\begin{proof}
		Due to the order of matchings for successive matchings $m,m+t$ for some $t\geq 1$, $C^*_{m} \leq C^*_{m+t}$. By definition, $B_{m+2} = \min(B_{m+1},C_{m+1})$ meaning that $B_{m+2} \leq B_{m + 1}$ so transitively $B_{m+t+1} \leq B_{m+1}$. Therefore
		\[B_{m+t+1} - C^*_{m+t}\leq B_{m+1} - C^*_m\implies \Delta_{m+t,\text{bounded}} \leq \Delta_{m,\text{bounded}}\]
	\end{proof}
	\begin{property}[Conditional optimality]
		\label{cond-monoton-optim}
		Assuming that $\Delta_m$, the unbounded effort, is monotonically non-decreasing as a function of $m$ (with $m$ in increasing order of $C^*_m$), ordered enumeration minimizes the total number of nodes searched across matchings.
	\end{property}
	\subsubsection{Approximate ordered enumeration}
	In the implementation of exhaustive ICTS, matchings are enumerated by taking the product of goal permutation indices as described above so they are not enumerated in increasing order of root SIC. To achieve this ordering, matchings would have to be exhaustively enumerated and stored in memory before being sorted. Storing all matchings requires $\mathcal{O}(M)$ memory which was found to be more problematic in practice than the time complexity of the sorting, $\mathcal{O}(M\log(M))$. This is why in implementation, ordered enumeration is approximated using a priority queue that maintains a subset of matchings of a fixed size $M_{pq}$\footnote{$M_{pq} = 10000$ was used in implementation but this could be much increased at the cost of memory}. Algorithm \ref{pq-algo} outlines the procedure.

	%Using the 
	%To generate matchings 
	%In the implementation of exhaustive matching, the algorithm used to generate matchings
	%The algorithm for generating 
	%In the implementation of ordered enumeration as described above, matchings are not generated in increasing order of root SIC. As a result, to achieve this ordering, matchings have to be exhaustively enumerated and stored in memory to be sorted. This requires $\mathcal{O}(M)$ memory which was found to be more problematic in practice than the time complexity of the sorting, $\mathcal{O}(M\log(M))$.

@@
